---
title: "String manipulation"
author: "NN"
date: "2021-10-25"
done: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

> This little tutorial aims to make you familiar with some of the functions of the `stringr` package and a few regular expressions. 

## Strings and escape sequences in R

##### Write a sentence with escape sequences.
Try the sentence: `"It's the end of the world!" he said.\` . 
Assign the string to a variable and try as`print()`, `cat()` and `writeLines()`.

```{r}
'"It\'s the end of the world!" he said.\\'

```

## `stringr` functions 

We will be using the `words` data that is built into `stringr`. 
The data set is available to you if you load the package.

```{r}

```


##### Select words that

1. Contain a `y`
```{r}

```

2. Start with `y`
```{r}

```


3. Contain a `y` within the word

```{r}
```

##### Extract the `y` and the previous character.

Note: Use the function `unique()` around the results to avoid printing many empty matches.

```{r}
```


##### Get the lengths of the first ten words

You can use `head(words, 10)` as a convenient way to access the elements of the `words` vector.
```{r}
```


## Viral research 

Read the genome sequence of the Hepatitis D virus: [hepd.fasta](https://biostat2.uni.lu/practicals/data/hepd.fasta).
For now, just execute the following:

```{r}
hepd_genome <- readr::read_lines("https://biostat2.uni.lu/practicals/data/hepd.fasta")
```

##### What is the length of the genome sequence? 

```{r}
```

##### What is the sequence composition? How often does each character occur?

```{r}
```


##### Find motifs in the sequence using `str_locate()`. 

Find all matches of the sequence $ATG$ in the genome sequence. 

```{r}
```


