  ---
title: "String manipulation"
subtitle: ""
author: "NN"
date: "23 Oct. 2022"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

> This little tutorial aims to make you familiar with some of the functions of the `stringr` package and a few regular expressions. 

## Strings and escape sequences in R

## `stringr` functions 

We will be using the `words` data that is built into `stringr`. 
The data set is available to you if you load the package.

```{r}
# Write your solution
```


##### Select words that contain a `y`

```{r}
# Write your solution
```

##### Retrieve a boolean vector that indicates which words start with `y`


```{r}
# Write your solution
```

##### Retrieve the indices for all words containing `ch` within the word. 

`ch` must not be present at the start or the end of the word. 

```{r}
# Write your solution
```

##### Extract the `y` and the previous character.

Note: Use the function `unique()` around the results to avoid printing many empty matches.

```{r}
# Write your solution
```


## Viral research 

Read the genome sequence of the Hepatitis D virus: [hepd.fasta](https://biostat2.uni.lu/practicals/data/hepd.fasta).
For now, just execute the following:

```{r}
hepd <- readr::read_lines("https://biostat2.uni.lu/practicals/data/hepd.fasta")
```

##### What is the length of the genome sequence?

```{r}
# Write your solution
```

##### What is the sequence composition? How often does each character occur?

```{r}
# Write your solution
```


##### Find motifs in the sequence 

Find the position of all matches of $ATG$ in the sequence. 

```{r}
# Write your solution
```


